---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data
Data were collected through monthly phone interviews. In conducting the BRFSS landline telephone survey, interviewers collect data from a randomly selected adult in a household. In conducting the cellular telephone version of the BRFSS questionnaire, interviewers collect data from an adult who participates by using a cellular telephone and resides in a private residence or college housing.

Because of the observational nature of the collection method, the scope of the inference is generalizable but cannot infer causation. Since only a percentage of people called completed the survey, there is also non-response bias in the results.

* * *

## Part 2: Research questions

**Research quesion 1:**
Do adults consuming more fruits and vegetables have more healthy days? Do female or male consume more fruits and vegetables?

**Research quesion 2:**
Do adults participated in 150+ minutes have longer hours of sleep than ones with 0 minutes?


**Research quesion 3:**
Do adults participated in 150+ minutes exercise consume more fruits and vegetabels and less sugar than ones with 0 minutes?

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}
q1 <- brfss2013 %>% select(X_frutsum, X_vegesum, physhlth, sex) %>% filter(!is.na(X_frutsum), !is.na(X_vegesum), !is.na(physhlth), (sex == "Male" | sex == "Female"))
q1 <- q1 %>% mutate(total_fv = X_frutsum + X_vegesum)
ggplot(data = q1, aes(x = physhlth, y = total_fv)) + geom_line()
q1 %>% group_by(sex) %>% summarise(mean_total_fv = mean(total_fv), sd_total_fv = sd(total_fv))
ggplot(data = q1, aes(x = sex, y = total_fv)) + geom_boxplot()
q1 %>% group_by(sex) %>% summarise(mean_total_fv = mean(total_fv), sd_total_fv = sd(total_fv), mean_hthdays = mean(physhlth), sd_hthdays = sd(physhlth))

```



**Research quesion 2:**

```{r}
q2 <- brfss2013 %>% select(X_pa150r2,sleptim1) %>% filter(!is.na(X_pa150r2), !is.na(sleptim1))
q2 %>% group_by(X_pa150r2) %>% summarise(mean_sleep_time = mean(sleptim1), sd = sd(sleptim1))

```



**Research quesion 3:**

```{r}
q3 <- brfss2013 %>% select(X_pa150r2, X_frutsum, X_vegesum, ssbsugar, ssbfrut2) %>%  filter(!is.na(X_pa150r2), !is.na(X_frutsum), !is.na(X_vegesum), !is.na(ssbsugar), !is.na(ssbfrut2))
q3 <- q3 %>% mutate(total_fv = X_frutsum + X_vegesum, total_sugar = ssbsugar + ssbfrut2)
q3 %>% group_by(X_pa150r2) %>% summarise(mean_fv = mean(total_fv), sd_fv = sd(total_fv), mean_sugar = mean(total_sugar), sd_sugar = sd(total_sugar))
```

